Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-16 WSSSPE1_report/WSSSPE1_summary_report.tex
Section 4: Minor changes.

The panel on developing and supporting software examined the challenges
around scientific software development and support, mainly focused on research
groups that also produce code in various forms. There was widespread agreement
that developing and maintaining software is hard, but best practices can help.
Several participants added that documentation is not just for users, and writing
API documentation, tutorials for building and deploying software, together with
documented development practices can be very helpful in bringing  new developers
into a project efficiently.

@@

Software is developed for many different purposes, and the requirements can
vary significantly depending on the intended audience. Most end-users make use
of either a graphical user interface of some kind, or a command line that may
offer input and output formats for running the code and analyzing its output(s).
When discussing backward compatibility it is these various interfaces that are
discussed. For software that builds on other software frameworks it is the
API (application programming interface) that is most important, and this can
encompass issues such as the source and binary interfaces to the software
libraries developed---with each potentially having a high maintenance cost if
they are to remain compatible over many years. When using command line programs
it is generally the command line switches as well as the input/output formats
that could incur significant costs if they are changed.

@@

Communities are extremely important in software projects, and both their
building and continued engagement need attention during the project life cycle.
Several of the submitted papers discussed how communities have been built around
projects, and what is needed to enable a project to grow. Among these are public
source code hosting, mailing lists, documentation, wikis, bug trackers, software
downloads, continuous integration, software quality dashboards, and of course,
a general web presence to tie all of these things together.

@@

The papers submitted to this panel included many general
recommendations for processes, practices, tools, etc. One of the papers
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a central common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center. The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g. Bitbucket or
GitHub. The centralized communication or point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.

@@

\subsection{Training Scientists to Develop and Support Software}
Part of the panel discussion focused around community structures and how academic communities are not taught how to evaluate cross-disciplinary work. 
One question raised was whether a software developer can be effective if they are not part
of the appropriate domain community, with responses that this depends on the 
specific problem and situation, and that ``T-shaped'' people who have both disciplinary depth as well as interdisciplinary and collaboration skills are important. 
The discussion focused on  whether we could teach software developers and domain scientists to 
collaborate together more effectively rather than trying to teach software developers
about domain science and domain scientists about software development practices. 
The end goal of this would be to have a single community with a spectrum of expertise across
domain science and software development rather than two separate communities.

>>>>

The panel on developing and supporting software examined the challenges
around scientific software development and support, mainly focused on research
groups that also produce code in various forms. There was widespread agreement
that developing and maintaining software is hard, but best practices can help.
Several participants added that documentation is not just for users, and writing
application programming interface (API) documentation, tutorials for building and deploying software, together with
documented development practices can be very helpful in bringing  new developers
into a project efficiently.

@@

Software is developed for many different purposes, and the requirements can
vary significantly depending on the intended audience. Most end-users make use
of either a graphical user interface of some kind, or a command line that may
offer input and output formats for running the code and analyzing its output(s).
When discussing backward compatibility it is these various interfaces that are
discussed. For software that builds on other software frameworks it is the
API that is most important, and this can
encompass issues such as the source and binary interfaces to the software
libraries developed---with each potentially having a high maintenance cost if
they are to remain compatible over many years. When using command-line programs
it is generally the command-line switches as well as the input/output formats
that could incur significant costs if they are changed.

@@

Communities are extremely important in software projects, and both their
building and continued engagement need attention during the project life cycle.
Several of the submitted papers discussed how communities have been built around
projects, and what is needed to enable a project to grow. Among these are 
public source-code hosting, mailing lists, documentation, wikis, bug trackers, software
downloads, continuous integration, software quality dashboards, and of course,
a general web presence to tie all of these things together.

@@

The papers submitted to this panel included many general
recommendations for processes, practices, tools, etc. One of the papers
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a central common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center. \note{What is software orphanage center?} The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g. Bitbucket or
GitHub. The centralized communication or point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.

@@

\subsection{Training Scientists to Develop and Support Software}
Part of the panel discussion focused around community structures and how academic communities are not taught how to evaluate cross-disciplinary work. 
One question raised was whether software developers can be effective if they are not part
of the appropriate domain community, with responses that this depends on the 
specific problem and situation, and that ``T-shaped'' people who have both disciplinary depth as well as interdisciplinary and collaboration skills are important. 
The discussion focused on  whether we could teach software developers and domain scientists to 
collaborate together more effectively rather than trying to teach software developers
about domain science and domain scientists about software development practices. 
The end goal of this would be to have a single community with a spectrum of expertise across
domain science and software development rather than two separate communities.
